---
title: "Data Analysis 01 Template"
author: "Student names here"
date: "3/2/2021"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, Nearly 250 women have been
fatally shot by police since 2015

(https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

# Load up the tidyverse
library(tidyverse)

```

## Load and Cleaning Data

[In this block of text, describe the source of the data, and write a basic data dictionary for data you are working with]

```{r}

# This data comes from a database compiled by The Washington Post of every fatal shooting in the U.S. by a police officer in the line of duty since Jan. 1, 2015.

# -------------------------------------------------------------------------------

# A basic data dictionary is as follows:

# "name" refers to the name of the victim who was fatally shot.
# "date" is the date of the shooting in YYYY-MM-DD format.
# "gender" is the victim's gender: Male, Female or Unknown
# "race" is the victim's race: W: White non-Hispanic, B: Black non-Hispanic, A: Asian, N: Native American, H: Hispanic, O: Other, None: unknown
# "city" and "state" are self-explanatory, though sometimes "city" contains a county if the specific city is unknown.
# "armed" refers to if a victim was armed with something police believed could cause harm. The cells indicate what weapon the victim possessed. "Unknown" means the victim had a weapon, but it is unknown what the weapon was. "Undetermined" means it is not known if a victim had a weapon. "Unarmed" means the victim had no weapon.
# "signs of mental illness" says if a victim had a history of mental illness based on news reports. TRUE means they did, FALSE means they did not.


# Load in WaPo's data as an object called fatal_police_shootings
fatal_police_shootings <- read_csv("data/fatal-police-shootings-data.csv")

# Clean required data and prepare for analysis if needed. 

glimpse(fatal_police_shootings)


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

## Eric

* **Sentence text**: “Since 2015, police have killed 26 unarmed women, including Taylor. Of those, 14 were White, seven were Black, four were Hispanic and one’s race was unknown. While about twice as many White women were shot and killed as Black women, White women account for five times the Black female population.”

* **Analysis summary**: Since this article was been written, two more unarmed women were shot and killed by police. So, the true count is 28 unarmed women, not 26. The two most recent deaths were Ashi Babbitt on Jan. 6, 2021 and Evatt Hernandez on Dec. 9, 2020. The race count in the above sentence is very close, but yes, since the article was published, it changed. There are still seven Black women and one woman whose race is unknown, but there are now 15 White women and five Hispanic women. The disparity they point out in the original sentence is still true given the disproportionate population difference between White women and Black women.

```{r}
# Put code to reverse engineer sentence here:

# Creates a new dataframe where it filters only for WOMEN who are UNARMED. Arranges them by most recent date of death.
unarmed_women_by_race <- fatal_police_shootings %>%
  filter(gender == "F" & armed == "unarmed") %>%
  arrange(desc(date))

# Creates new dataframe pulling from the dataframe made above. Uses count() function to get a count of how many of each race have died. sort = TRUE displays this count from largest to smallest

counts_of_race <- unarmed_women_by_race %>%
  group_by(race) %>%
  count() %>%
  arrange(desc(n))
  #count(race, race, wt = NULL, sort = TRUE, name = NULL)
  


# Display results of code below this codeblock
unarmed_women_by_race

counts_of_race

```

### Sentence 2

## Eric

* **Sentence text**: “Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known. But within this small subset, Black women, who are 13 percent of the female population, account for 20 percent of the women shot and killed and 28 percent of the unarmed deaths.”

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here


# Step 1: Get % of Black women compared to all people in the dataset

# Step 2: Get % of Black women compared to all women in the dataset

# Step 3: Get % of Black unarmed women compared to all unarmed people in the dataset.



# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: “About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed.”

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

## Pierce

* **Sentence text**: “Since The Washington Post began tracking fatal shootings by police in 2015, officers have fatally shot 247 women out of the more than 5,600 people killed overall.”

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: “The starkest difference between women and men is the rate: Women account for about half the population, but 4 percent of the killings.”

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-